---
title: "Mercado_CID_2024"
author: "Reinaldo Mercado"
date: "2024-July"
output: html_document
---

These are code used to conduct the statistic and create the tables and figures 
for the following manuscript:

Mercado-Hernandez et al.
Obesity is associated with increased pediatric dengue virus infection and disease: 
A 9-year cohort study in Managua, Nicaragua. 
Clinical Infectious Diseases. 2024. ciae3602024. 

NOTE: 
This is an incomplete version. 
I am currently cleaning and annotating the code for easier reading. 
I will update this script as soon as possible. 

 
START

Libraries
```{r Text}
library(dplyr)
library(AGD) #step 1 - calculate BMI z-scores
library(epiR) #Step 2 - 
library(gtsummary) #Step 2, 3 and 5 - Tabulating results
library(geepack) #Step 4 - logistic regression with GEE
library(forester) #Step 6 - Plotting results
```


1- Preparing data for analysis
```{r Data}
####1 - Data####
#1.1 - Load data
load("Data.Rdata")
#_______________________________________________________________________________
#1.2 - BMIz Z-score (BMIz) category 
A_Data$BMI_Z_score <- 
  AGD::y2z(
    y=A_Data$BMI, 
    x=A_Data$Age, 
    sex=A_Data$Sex, 
    ref = who.bmi, 
    tail.adjust=T)
#_______________________________________________________________________________
#1.3 - BMIz categories 

#_#1.3.1 Calculate BMIz
A_Data <- A_Data %>% 
  mutate(BMIz_Category = 
#__#BMIz for children 2 to 5 years old):
#__#__#__#Underweight: < -2SD.
#__#__#__#Normal weight: >=-2 to <=+2SD.
#__#__#__#Overweight: >+2SD to <=+3SD. 
#__#__#__#Obese: >+3SD.
         ifelse(HW_Age_Year < 5 & BMI_Z_score < -2, "Underweight", 
         ifelse(HW_Age_Year < 5 & BMI_Z_score >= -2 & BMI_Z_score <= 2, "Normalweight", 
         ifelse(HW_Age_Year < 5 & BMI_Z_score > 2 & BMI_Z_score <=3, "Overweight",
         ifelse(HW_Age_Year < 5 & BMI_Z_score > 3, "Obese",
#__#BMIz for children 5 to 18 years old):
#__#__#__#Underweight: < -2SD.
#__#__#__#Normal weight: >=-2 to <=+1SD
#__#__#__#Overweight: >+1SD to <=+2SD (equivalent to BMI 25 kg/m2 at 19 years)
#__#__#__#Obese: >+2SD (equivalent to BMI 30 kg/m2 at 19 years)
         ifelse(HW_Age_Year >= 5 & BMI_Z_score < -2, "Underweight", 
         ifelse(HW_Age_Year >= 5 & BMI_Z_score >= -2 & BMI_Z_score <= 1, "Normalweight", 
         ifelse(HW_Age_Year >= 5 & BMI_Z_score > 1 & BMI_Z_score <=2, "Overweight",
         ifelse(HW_Age_Year >= 5 & BMI_Z_score > 2, "Obese", NA)))))))))

#_#1.3.2 Make BMIz_Category = "NA" into "No_data"
A_Data$BMIz_Category <- 
  replace(A_Data$BMIz_Category,is.na(A_Data$BMIz_Category),"No_data")

#_______________________________________________________________________________
#1.4 - Only include participants with recorded BMIz Z-score category data
A_Data <- A_Data %>% 
  filter(BMIz_Category == "Normalweight" | 
           BMIz_Category == "Underweight" |
           BMIz_Category == "Overweight" |
           BMIz_Category == "Obese") 
#_______________________________________________________________________________
#1.5 - Variable order
#_#1.5.1 BMIz_Category
A_Data$BMIz_Category <- factor(
      A_Data$BMIz_Category, 
      levels = c("Normalweight", "Underweight", "Overweight", "Obese"))
#_#1.5.1 DENV_iELISA_Titer
A_Data$DENV_iELISA_Titer <- factor(
      A_Data$DENV_iELISA_Titer, 
      levels = c(
      "[0-10)", "[10-20)", "[20-80)", "[80-320)", "[320-1280]", ">1280"))
#_#1.5.1 Sex
A_Data$Sex <- factor(
      A_Data$Sex, 
      levels = c(
      "Female", "Male"))
#_______________________________________________________________________________
#1.6 - Add number of observations per Code_ID
#List of Code_ID with observation counts
B_obs <- A_Data %>% count(Code_ID) %>% rename(No_obs = "n")
#Merge to Cohort data
A_Data <- left_join(
  A_Data, B_obs, by = "Code_ID")
#_______________________________________________________________________________
####2 - Table 1####
#Coming soon
#I am working on cleaning and annotating these code for easier understanding
#They will be uploaded in the next version of this file. 
#_______________________________________________________________________________
####3 - Table 2####
#Coming soon
#I am working on cleaning and annotating these code for easier understanding
#They will be uploaded in the next version of this file. 
#_______________________________________________________________________________
####4 - Logistic regression GEE####
#_#4.1 Bivariable models
C_GEE_1_Bi <- geeglm(Outcome ~ BMIz_Category, 
            data = A_Data_1,
            id = Code_ID, family = binomial, corstr = "exchangeable")
#_#4.2 Multivariable_B
C_GEE_1_Mul <- geeglm(Outcome ~ BMIz_Category + 
                        Age + Sex + DENV_iELISA_Titer, 
            data = (A_Data_1 %>% 
                      filter(DENV_iELISA_Titer != "Missing")), 
            id = Code_ID, family = binomial, corstr = "exchangeable")
#Note: The same code was followed for each of the tested outcomes
#Note: For 1) Dengue given DENV and 2) dengue clinical manifestations
#analysis, the following filter was used: 
#A_Data_1 %>% filter(DENV_infection == "TRUE")
#_______________________________________________________________________________
####5 - Models results into table####
#_#5.1 Functions for extracting models results
#_#5.1.1 Bivariable
Model_table_Bi <- function(data, exponentiate, digits_est, digits_pvalue) {
  tbl_regression(data,
                 exponentiate = exponentiate,
                 label = list(
                 BMIz_Category ~ "BMIz Category"),
                 estimate_fun = purrr::partial(style_ratio, digits = digits_est), 
                 pvalue_fun = purrr::partial(style_sigfig, digits = digits_pvalue)) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_p(t = 0.05, q = FALSE) %>%
    bold_labels() %>%
    italicize_levels()
}
#_______________________________________________________________________________
#_#5.1.2 Multivariate
Model_table_Multi <- function(data, exponentiate, digits_est, digits_pvalue) {
  tbl_regression(data,
                 exponentiate = exponentiate,
                 label = list(
                BMIz_Category ~ "BMIz Category",
                Age ~ "Age", 
                Sex ~ "Sex", 
                DENV_iELISA_Titer ~ "DENV antibody titer"),
                 estimate_fun = purrr::partial(style_ratio, digits = digits_est), 
                 pvalue_fun = purrr::partial(style_sigfig, digits = digits_pvalue)) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_p(t = 0.05, q = FALSE) %>%
    bold_labels() %>%
    italicize_levels()
}
#_______________________________________________________________________________
#_#5.2 Making the tables
#_#5.2.1 Bivariable
D_Table <- Model_table_Bi(C_GEE_1_Bi, TRUE, 2, 4)
#_#5.2.2 Multivariate
D_GEE_1B_df <- Model_table_Multi(C_GEE_1_Mul, TRUE, 2, 4)
#Note: The same code was followed for each of the tested outcomes
#_______________________________________________________________________________
####6 - Models results into forest plots####
#Coming soon
#I am working on cleaning and annotating these code for easier understanding
#They will be uploaded in the next version of this file. 
#_______________________________________________________________________________
####END####
```
END